Spin canonical divisors of log stacky curves

### 2015-06-24 stacky_write_up/spin_cring.tex
minor changes section 3
replace all ram pt -> stacky pt, ram ord -> stab ord

In Voight and Zureick-Brown, \cite{vzb:stacky} examine a much more general setting, generalizing Petri's theorem in many directions.
A divisor $\Delta \in \d X$ is a {\bf log divisor} if $\Delta = \sum_{i}^{}P_i \in \di X,$ with all $P_i$ distinct. Let $(\sx,\Delta)$ denote a tame seperably rooted stacky curve over a field $k$ so that $X$ is the coarse space of $\sx$, as defined in \cite{vzb:stacky}[Section 5.2] so that $\di \sx = \div X \bigoplus_{i=1}^r \langle \frac{1}{e_i}P_i \rangle,$ with $e_i \in \BZ$. Let $K_\sx = K_X + \sum_{i=1}^{r}\frac{e_i-1}{e_i}P_i.$ In this case, we say $\sx$ has {\bf ramification order} $e_i$ at $P_i$. Call $\halfcan \in \di \sx$ a {\bf log half-canonical divisor} on $(\sx,\Delta)$, be a divisor $\halfcan \in \di \sx$ so that $2\halfcan \sim K_\sx + \Delta.$ If $X$ has genus $g$ and $\delta = \deg \Delta,$ then we say $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$. A {\bf log spin curve} is a triple $(X,\Delta,\halfcan)$ where $(X,\Delta)$ and $\halfcan$ is a log half-canonical divisor on $(\sx,\Delta)$. The central object of study in this paper is the {\bf log spin canonical ring} of $(X,\Delta,L)$ is
\begin{align*}
	R(X,\Delta,L) = \oplus_{n=0}^\infty H^0(X,nL)
\end{align*}
The main theorem of this paper is the following bound on the degree of generators and relations of arbitrary log spin canonical rings:

@@

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3,
		4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ and $y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4, y_5]$ with $grlex$ so that $y_5 \prec 
		y_4 \prec y_3$
		and equip the ring $k[y_3, y_4, y_5, x]$ with the block 
		order so that $R = k[y_3, y_4, y_5, x]/I$. Let,
		\begin{align*}
			J &= gin_\prec(I')k[x,y_3,y_4,y_5] \\
			&+\langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
			&+\langle y_iy_k \mid 4 \leq i \leq j \leq 5.\rangle 
		\end{align*}
			Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{lem}

@@

\ssec{Inductive Lemma: Raising Ramification Orders}
\label{ssec:ram_orders}
In this section, we aim to prove the inductive theorem, Theorem
~\ref{thm:ram_order_ind}. Under fairly general
conditions, this theorem will allow us to show that if Theorem ~\ref
{thm:ram_order_ind} holds for a curve with signature
$(0;e_1',\ldots, e_r';0)$ then it also holds for a curve with
signature $(0;e_1, \ldots, e_r; 0)$ with $e_i' - e_i \in \{0,2\}$.

@@

\ssec{Base Cases}
\label{ssec:g_0_base}
Now we have an inductive statement on the number of ramified
points in Lemma ~\ref{lem:sat_three_induction} and
an inductive statement on ramification orders in Theorem
~\ref{thm:ram_order_ind}. After verifying that explicit pairs
$(\sx, J)$ of stacky curves $\sx$ with particular signatures and
particular choices of $J$ are admissible, the application of the
two inductive statements demonstrate that the spin canonical ring
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1,
\ldots, e_r)$ with relations generated in degree at most $2e$ for
$\sx$ with any signature $\sigma = (0; e_1, \ldots , e_r; 0)$ outside
of a finite set of exceptional cases.

@@

Through the application of two processes of induction, by adding 
points via Lemma ~\ref{lem:sat_three_induction} and by raising
ramification orders via Theorem ~\ref{thm:ram_order_ind},
we obtain Theorem ~\ref{thm:g_0_generators_relations} for any tame
log spin curve $(\sx, \Delta, L)$ over a perfect field $k$ with
signature $\sigma$ covered by this induction process.
In particular, this induction process covers all signatures $\sigma
= (0; e_1, \ldots, e_r; 0)$ such that either $r \geq 5$ or
for some listed base case $(\sigma', J)$ with $\sigma' = (0; e_1',
\ldots, e_r'; 0)$, $e_i \geq e_i'$ for all $i$.
Before we prove the main theorem, we describe the cases not be
covered by induction.

@@

\begin{proof}
If $\sx$ has no ramification points, then we have the classical
case which is done in Chapter 4 of Voight and Zureick-Brown
\cite{vzb:stacky}.

@@

If $r > 5$, then we can use Lemma
~\ref{lem:sat_three_induction} to add ramified points with
ramification order $3$ to case (k) of Lemma
~\ref{lem:g_0_admissible_cases}, which corresponds to $\sigma_{11}
= (0; 3, 3, 3, 3, 3; 0), J_{11} = \{1, 2, 3, 4, 5\}$. This case
clearly satisfies the conditions of Lemma
~\ref{lem:sat_three_induction} (recall from Subsection ~\ref
{ssec:g_0_saturation} that saturation is 3 here), and the
immediate consequence of parts (a) and (c) of Lemma
~\ref{lem:sat_three_induction} is that any $R(\sx', \Delta,
\halfcan')$ corresponding to signatures $\sigma'$ with ramification
orders all equal to $3$ for any $r > 5$ is generated up to
degree $e := \max(5, e_1', \ldots, e_r')$ with relations generated
up to degree $2e$. Furthermore, Lemma ~\ref{lem:n-3s-admissible}
implies that these cases satisfy all of the conditions of Theorem
~\ref{thm:ram_order_ind}.

So if $r > 5$, then $\sigma$ ``lies above'' the case $(\sigma', J)$
where $\sigma'$ has $r$ many ramification points all of order $3$
and $J = \{1, \ldots, r\}$ and these underlying cases satisfy
all of the hypotheses of Theorem ~\ref{thm:ram_order_ind}.
If $r \leq 5$, then $\sigma$ ``lies above'' one of the base cases
of Lemma ~\ref{lem:g_0_admissible_cases}, all of which satisfy
the hypotheses of Theorem ~\ref{thm:ram_order_ind} as well.

So for any $r > 0$, there is a $(\sigma', J)$, either from
Lemma ~\ref{lem:g_0_admissible_cases} or corresponding to
the $r > 5$ case with ramification orders all $3$, with $\sigma' =
(0; e_1', \ldots, e_r'; 0)$ satisfying the property that for all $i
\in J$, $e_i \geq e_i'$ and $e_i = e_i'$ otherwise.

@@

%By Lemma ~\ref{lem:g_0_admissible_cases}, we know that for
%any tame log spin curves $(\sx', \Delta, \halfcan')$ over a perfect
%field $k$ with signature $\sigma'$ and set $J$ such that $(\sigma',
%J)$ is one of the listed base cases, $(\sigma', J)$ is admissible.
%Furthermore, $R' := R(\sx', \Delta, \halfcan')$ is generated as a $k
%$- algebra by elements of degree at most $e = \max(5, e_1, \ldots, e
%_r)$ with relations in degree at most $2e$. It also admits a
%presentation
%
%\begin{align*}
	%R' \cong \left( k[x_1, \ldots, x_m] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
%\end{align*}
%
%\noindent
%given by admissibility satisfying the statements that for all $i \in 
%J$, $e_i > \deg z$ for any generator $z$ of $R'$ and that $in_\prec
%(I')$ is minimally generated by products of two monomials under
%some ordering.
%
%Then we can apply two processes of induction, by adding points via
%Lemma ~\ref{lem:sat_two_induction} and by raising
%ramification orders via Theorem ~\ref{thm:ram_order_ind},
%to obtain the desired result for any tame log spin curve $(\sx,
%\Delta, L)$ over a perfect field $k$ with signature $\sigma$
%covered by this induction process.
%
%
%
%A process of induction raising the ramification orders of
%the ramified points of the spin curves corresponding to Lemma
%~\ref{lem:g_0_admissible_cases} cases (a)-(c) via Theorem
%~\ref{thm:ram_order_ind} will demonstrate the desired generator and
%relation degree results for curves with three ramified points.
%Similarly, this induction process on Lemma
%~\ref{lem:g_0_admissible_cases} case (d) will give the desired
%result for curves with four ramified points excluding signature
%$(0; 3, 3, 3, 3; 0)$. A process of induction increasing the
%number of ramified points via Lemma
%~\ref{lem:sat_three_induction} on Lemma ~\ref{lem:g_0_admissible_cases}
%case (e) combined with the induction process on ramification orders
%will yield the result for curves with at least five ramified points.
\end{proof}

>>>>

In Voight and Zureick-Brown, \cite{vzb:stacky} examine a much more general setting, generalizing Petri's theorem in many directions.
A divisor $\Delta \in \d X$ is a {\bf log divisor} if $\Delta = \sum_{i}^{}P_i \in \di X,$ with all $P_i$ distinct. Let $(\sx,\Delta)$ denote a tame seperably rooted stacky curve over a field $k$ so that $X$ is the coarse space of $\sx$, as defined in \cite{vzb:stacky}[Section 5.2] so that $\di \sx = \div X \bigoplus_{i=1}^r \langle \frac{1}{e_i}P_i \rangle,$ with $e_i \in \BZ$. Let $K_\sx = K_X + \sum_{i=1}^{r}\frac{e_i-1}{e_i}P_i.$ In this case, we say $\sx$ has {\bf stabilizer order} $e_i$ at $P_i$. Call $\halfcan \in \di \sx$ a {\bf log half-canonical divisor} on $(\sx,\Delta)$, be a divisor $\halfcan \in \di \sx$ so that $2\halfcan \sim K_\sx + \Delta.$ If $X$ has genus $g$ and $\delta = \deg \Delta,$ then we say $\sx$ has signature $(g;e_1,\ldots, e_r;\delta)$. A {\bf log spin curve} is a triple $(X,\Delta,\halfcan)$ where $(X,\Delta)$ and $\halfcan$ is a log half-canonical divisor on $(\sx,\Delta)$. The central object of study in this paper is the {\bf log spin canonical ring} of $(X,\Delta,L)$ is
\begin{align*}
	R(X,\Delta,L) = \oplus_{n=0}^\infty H^0(X,nL)
\end{align*}
The main theorem of this paper is the following bound on the degree of generators and relations of arbitrary log spin canonical rings:

@@

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3,
		4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ and $y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4, y_5]$ with $grlex$ so that $y_5 \prec 
		y_4 \prec y_3$
		and equip the ring $k[y_3, y_4, y_5, x]$ with the block 
		order so that $R = k[y_3, y_4, y_5, x]/I$. Let,
		\begin{align*}
			\initial_\prec(I) &= \initial_\prec(I)k[x,y_3,y_4,y_5] \\
			&+\langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
			&+\langle y_iy_k \mid 4 \leq i \leq j \leq 5.\rangle 
		\end{align*}
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{lem}

@@

\ssec{Inductive Lemma: Raising stabilizer orders}
\label{ssec:ram_orders}
In this section, we aim to prove the inductive theorem, Theorem
~\ref{thm:ram_order_ind}. Under fairly general
conditions, this theorem will allow us to show that if Theorem ~\ref
{thm:ram_order_ind} holds for a curve with signature
$(0;e_1',\ldots, e_r';0)$ then it also holds for a curve with
signature $(0;e_1, \ldots, e_r; 0)$ with $e_i' - e_i \in \{0,2\}$.

@@

\ssec{Base Cases}
\label{ssec:g_0_base}
Now we have an inductive statement on the number of ramified
points in Lemma ~\ref{lem:sat_three_induction} and
an inductive statement on stabilizer orders in Theorem
~\ref{thm:ram_order_ind}. After verifying that explicit pairs
$(\sx, J)$ of stacky curves $\sx$ with particular signatures and
particular choices of $J$ are admissible, the application of the
two inductive statements demonstrate that the spin canonical ring
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1,
\ldots, e_r)$ with relations generated in degree at most $2e$ for
$\sx$ with any signature $\sigma = (0; e_1, \ldots , e_r; 0)$ outside
of a finite set of exceptional cases.

@@

Through the application of two processes of induction, by adding 
points via Lemma ~\ref{lem:sat_three_induction} and by raising
stabilizer orders via Theorem ~\ref{thm:ram_order_ind},
we obtain Theorem ~\ref{thm:g_0_generators_relations} for any tame
log spin curve $(\sx, \Delta, L)$ over a perfect field $k$ with
signature $\sigma$ covered by this induction process.
In particular, this induction process covers all signatures $\sigma
= (0; e_1, \ldots, e_r; 0)$ such that either $r \geq 5$ or
for some listed base case $(\sigma', J)$ with $\sigma' = (0; e_1',
\ldots, e_r'; 0)$, $e_i \geq e_i'$ for all $i$.
Before we prove the main theorem, we describe the cases not be
covered by induction.

@@

\begin{proof}
If $\sx$ has no stacky points, then we have the classical
case which is done in Chapter 4 of Voight and Zureick-Brown
\cite{vzb:stacky}.

@@

If $r > 5$, then we can use Lemma
~\ref{lem:sat_three_induction} to add ramified points with
stabilizer order $3$ to case (k) of Lemma
~\ref{lem:g_0_admissible_cases}, which corresponds to $\sigma_{11}
= (0; 3, 3, 3, 3, 3; 0), J_{11} = \{1, 2, 3, 4, 5\}$. This case
clearly satisfies the conditions of Lemma
~\ref{lem:sat_three_induction} (recall from Subsection ~\ref
{ssec:g_0_saturation} that saturation is 3 here), and the
immediate consequence of parts (a) and (c) of Lemma
~\ref{lem:sat_three_induction} is that any $R(\sx', \Delta,
\halfcan')$ corresponding to signatures $\sigma'$ with ramification
orders all equal to $3$ for any $r > 5$ is generated up to
degree $e := \max(5, e_1', \ldots, e_r')$ with relations generated
up to degree $2e$. Furthermore, Lemma ~\ref{lem:n-3s-admissible}
implies that these cases satisfy all of the conditions of Theorem
~\ref{thm:ram_order_ind}.

So if $r > 5$, then $\sigma$ ``lies above'' the case $(\sigma', J)$
where $\sigma'$ has $r$ many stacky points all of order $3$
and $J = \{1, \ldots, r\}$ and these underlying cases satisfy
all of the hypotheses of Theorem ~\ref{thm:ram_order_ind}.
If $r \leq 5$, then $\sigma$ ``lies above'' one of the base cases
of Lemma ~\ref{lem:g_0_admissible_cases}, all of which satisfy
the hypotheses of Theorem ~\ref{thm:ram_order_ind} as well.

So for any $r > 0$, there is a $(\sigma', J)$, either from
Lemma ~\ref{lem:g_0_admissible_cases} or corresponding to
the $r > 5$ case with stabilizer orders all $3$, with $\sigma' =
(0; e_1', \ldots, e_r'; 0)$ satisfying the property that for all $i
\in J$, $e_i \geq e_i'$ and $e_i = e_i'$ otherwise.

@@

%By Lemma ~\ref{lem:g_0_admissible_cases}, we know that for
%any tame log spin curves $(\sx', \Delta, \halfcan')$ over a perfect
%field $k$ with signature $\sigma'$ and set $J$ such that $(\sigma',
%J)$ is one of the listed base cases, $(\sigma', J)$ is admissible.
%Furthermore, $R' := R(\sx', \Delta, \halfcan')$ is generated as a $k
%$- algebra by elements of degree at most $e = \max(5, e_1, \ldots, e
%_r)$ with relations in degree at most $2e$. It also admits a
%presentation
%
%\begin{align*}
	%R' \cong \left( k[x_1, \ldots, x_m] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
%\end{align*}
%
%\noindent
%given by admissibility satisfying the statements that for all $i \in 
%J$, $e_i > \deg z$ for any generator $z$ of $R'$ and that $in_\prec
%(I')$ is minimally generated by products of two monomials under
%some ordering.
%
%Then we can apply two processes of induction, by adding points via
%Lemma ~\ref{lem:sat_two_induction} and by raising
%stabilizer orders via Theorem ~\ref{thm:ram_order_ind},
%to obtain the desired result for any tame log spin curve $(\sx,
%\Delta, L)$ over a perfect field $k$ with signature $\sigma$
%covered by this induction process.
%
%
%
%A process of induction raising the stabilizer orders of
%the ramified points of the spin curves corresponding to Lemma
%~\ref{lem:g_0_admissible_cases} cases (a)-(c) via Theorem
%~\ref{thm:ram_order_ind} will demonstrate the desired generator and
%relation degree results for curves with three ramified points.
%Similarly, this induction process on Lemma
%~\ref{lem:g_0_admissible_cases} case (d) will give the desired
%result for curves with four ramified points excluding signature
%$(0; 3, 3, 3, 3; 0)$. A process of induction increasing the
%number of ramified points via Lemma
%~\ref{lem:sat_three_induction} on Lemma ~\ref{lem:g_0_admissible_cases}
%case (e) combined with the induction process on stabilizer orders
%will yield the result for curves with at least five ramified points.
\end{proof}
